DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-02 paper/evaluation.tex
Add preliminary conclusion and revised evaluation
Also added redrawn Figure 1 powerpoint and image.

In evaluating the real-time efficacy of the Raspberry Pi 3, the same methodology is utilized in all 
experiments conducted. The performance of the Pi is measured over a set of 1001 video frames that are 
each individually fed to the model. The processing time for the first frame is ignored as, due to 
cache warmup, it is uncharacteristically high and doesn't accurately represent the Pi's capabilities. A 
deadline of 50 ms, or 20 Hz, is used as a baseline to assess the Pi's ability to complete all 
necessary real-time operations in a timely manner. Also, please note that frame processing times 
would be approximately the same if the input stream was a 
camera instead of a video.

\subsection{Real-Time Operations}
In a real-time operations, the system has to be capable of consistently executing all necessary 
functions before their given deadlines. In the case of our platform, it has to process every given frame 
and get the predicted angles from the model. In order to determine if the Raspberry Pi 3 is capable of 
this, we tested the platform and recorded the time it took for the Pi to complete all real-time 
functions for every given frame.For this experiment, all four of the Pi's cpu cores were utilized, and 
only one model was run. As can be seen in Figure 1, the Pi was able to meet almost all of its deadlines, 
while only missing it in a few instances.

@@

In our platform, three main real-time tasks are performed during autonomous operation. 
In order, those operations are: (1) capturing and reading the input frame from the designated camera 
or video stream, (2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) 
feeding the frame to, and getting the angle prediction from, the model. However, the time it takes to 
complete each of the operations will differ, with at least one of them being the dominating step in 
processing each frame.

\begin{table}[h]
  \centering
  \begin{tabular} {| l | l | l | l | l |}
    \hline
    \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
    Frame Capture & 2.28 & 4.94 & 63.31 &  .51\\ \hline
    Preprocessing & 3.09 & 4.6 & 3.31 & .1 \\ \hline
    Angle Prediction & 37.3 & 51.03 & 45.48 & 2.75 \\ 
    \hline
  \end{tabular}
  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
\end{table}

In order to determine which operation(s) take the longest to execute, we measured the time it 
took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
and only one model was run. As is shown in Fig. 1, the angle prediction operation (3) consumes the 
majority of the processing for each frame. Furthermore, the time it takes for the operation to 
complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
other hand, both the frame capture (1) and preprocessing (2) operations take substantially less time 
and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.

@@

\begin{table}[h]
  \centering
  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
    \hline
    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
    1 & 61.96 & 66 & 63.31 &  .51\\ \hline
    2 & 50.49 & 71.55 & 70.03 & 4.18 \\ \hline
    3 & 48.11 & 72.22 & 58.45 & 2.8 \\ \hline
    4 & 42.67 & 56.37 & 50.70 & 2.8 \\
    \hline
  \end{tabular}
  \caption{Real-time performance of the Raspberry Pi 3 depending on the number of cores used.}
\end{table}

As is depicted in Table 1, the Raspberry Pi performed better, on average, when it utilized more 
cores. With 4 cores, the Pi was able to meet the vast majority of its 50 ms deadlines, doing so in 
almost 99\% of the processed frames. The Pi had the worst average performance when using only 1 core, 
as it was unable to meet any of the 50 ms deadlines (its fastest processing time was just under 60 
ms). Another observation is that the difference between using 2 cores and 3 cores is relatively 
small. On average, using 3 cores only performed better by 2 ms, so the addition of one core in that 
specific case offers relatively little improvement. However, please note that the average time when 
using 3 cores was below the 50 ms deadline, while the average time using 2 cores was slightly above it. 
One important observation that can be made is that of consistency, as using only 1 core had a much 
steadier performance. As a result, the use of multiple cores is very beneficial in terms of reducing 
the time it takes to complete real-time operations, but may ultimately result in processing times 
that are more volatile.

\subsection{Multimodel Performance}
In a real world scenario, it is highly probable that multiple models will need to be used 
simultaneously in order to perform various important tasks (angle prediction, object detection, 
etc.)\cite{}. As such, we also tested the capability of the Raspberry Pi to run multple models at the 
same time, and measure whether all models are able to meet their respective deadlines on a consistent 
basis. Specifically, the Pi is tested in the cases of running 2 and 4 models simultaneously. For each 
case, all models are allocated an equal number of cores, with 2 models having 2 cores each, and 4 
models having 1 core each.

\begin{table*}
  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
  \hline
  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
  1 & 0,1 & 51.35 & 3.04E+10 & 4.78E+08 & 1.58 & 3.31E+09 & 3.68E+08 & 11.12\\ \hline
  2 & 0,1 & 58.03 & 3.04E+10 & 4.91E+08 & 1.61 & 3.91E+09 & 4.26E+08 & 10.88 \\ \hline
  2 & 2,3 & 56.4 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
  \end{tabular}
  \caption{Real-time performance of the Raspberry Pi 3 when 2 models are running simultaneously.}
\end{table*}

Ideally, each model run would replicate a single model run. However, Table 2 shows that such was not 
the case. In the tests where 2 models ran simultaneously, both of the models showed average time 
increases of around 5-7 ms, around 10\%, when compared to a baseline of 1 model running on two cores. 
This change, however, did not seem to be caused by any form of cache interference as both L1 and L2 
cache misses remained constant, regardless of the number of models running. The number of L1 misses 
being close to 1.6\% of all references and the number of L2 misses being about 11\% of all 
references.

\begin{table*}
  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
  \hline
  \textbf{num models} & \textbf{core} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
    misses} & \textbf{L1 miss \%} & textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
    1 & 0 &  62.48 & 2.78E+10 & 4.36E+08 & 1.57 & 2.83E+09 & 3.59E+08 & 12.68  \\ \hline
    4 & 0 & 77.9 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
    4 & 1 & 78.89 & 2.79E+10 & 4.64E+08 & 1.67 & 3.42E+09 & 4.38E+08 & 12.82 \\ \hline
    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.6 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
    4 & 3 & 77.87 & 2.79E+10 & 4.45E+08 & 1.60 & 3.41E+09 & 4.39E+08 & 12.88 \\ \hline
  \end{tabular}
  \caption{Real-time performance of the Raspberry Pi when 4 models are running simultaneously.}
\end{table*}

The difference was even greater in the case of 4 models running concurrently, as each one displayed 
an average time increase of approximately 15 ms, around 30\%, when compared to a single model 
running on 1 core. Once again, though, cache interference was not a factor as the number of L1 cache 
misses remained 1.6\% of all references, and the number of L2 cache misses stayed at 13\% of all 
references.

@@

The most notable issue that can affect the cpu clock speed is that of the power supplied to the 
Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments done with a power supply that 
only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
between operating at 600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly 
recommended, that the power supply used for the Raspberry Pi 3 be capable of outputting 2 Amps, 
otherwise optimal performance isn't guaranteed.

Another factor that can affect clock speed is that of the cpu's temperature. Some model operations can 
be computationally intensive, thus it is possible for the temperature of the cpu to become relatively 
high. This can be especially problematic in situations where multiple models are running 
simultaneously on the Pi. As a consequence, thermal throttling may be used to decrease the clock 
speed so that the cpu temperature stays at a safe level. As such, the Raspberry Pi may not be suited 
for prolonged use, especially in cases where the workload is relatively larger, like running multiple 
models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
off or made idle so that the cpu is given time to cool down.

>>>>

In evaluating the real-time efficacy of the DeepPicar, the methodology is consistent across all 
experiments. The performance of the platform is measured over a set of 1001 video frames that are 
each individually fed to the model. The processing time for the first frame is omitted as, due to 
cache warmup, it is uncharacteristically high and doesn't accurately represent the Picar's capabilities. 
A WCET of 50 ms, or 20 Hz, is used as a baseline to assess the platform's ability to complete all 
necessary real-time operations in a timely manner.

\subsection{Real-Time Operations}
We seek to determine if the DeepPicar is capable of consistently executing all necessary functions 
before their given deadlines. In the case of our platform, it has to process every given frame and get 
the predicted angles from the model within a WCET of 50 ms. In order to determine if this was achievable, 
we tested the platform by running a single model that utilized all 4 cpu cores and measured the time it 
took for each frame to be processed. The performance of the DeepPicar can be seen in Figure 6. We found 
that DeepPicar was able to completely process the vast majority of the provided frames within 50 ms, and 
was unable to do so for very few frames.

@@

In our platform, three main real-time tasks are performed during autonomous operation. In order, these 
operations are: (1) capturing and reading the input frame from the designated camera or video stream, 
(2) preprocessing the acquired frame so that it is compatible with the DNN, and (3) feeding the frame 
to, and getting the angle prediction from, the model. We aim to determine which operation(s), if any, 
require the most time to be performed.

\begin{table}[h]
  \centering
  \begin{tabular} {| l | l | l | l | l |}
    \hline
    \textbf{operation} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
    Frame Capture & 2.28 & 4.94 & 4.54  & 0.52\\ \hline
    Preprocessing & 3.09 & 4.60 & 3.31 & 0.10 \\ \hline
    Angle Prediction & 37.30 & 51.03 & 45.48 & 2.75 \\  \hline
    Total Time & 42.67 & 56.37 & 50.70 & 2.80 \\
    \hline
  \end{tabular}
  \caption{Real-time performance of DeepPicar for each autonomous vehicle operation performed.}
\end{table}

In order to determine which operation(s) take the longest to execute, we measured the time it 
took for each step to be completed. For this experiment, all four of the Pi's cpu cores were utilized, 
and only one model was run. As is shown in Table II, the angle prediction operation consumes the 
majority of the processing for each frame. Furthermore, the time it takes for the operation to 
complete is volatile, and can range anywhere between 30 ms and 50 ms for any particular frame. On the 
other hand, both the frame capture and preprocessing operations take substantially less time 
and are relatively more consistent in their times, at 2 ms and 3 ms, respectively.

@@

\begin{table}[h]
  \centering
  \begin{tabular} {| l | l | l | l | l | l | l | l | l | l |}
    \hline
    \textbf{num cores} & \textbf{mean} & \textbf{max} & \textbf{99pct} & \textbf{stdev} \\ \hline 
    1 & 61.96 & 66.00 & 63.31 & 0.51\\ \hline
    2 & 50.49 & 71.55 & 70.03 & 4.18 \\ \hline
    3 & 48.11 & 72.22 & 58.45 & 2.80 \\ \hline
    4 & 42.67 & 56.37 & 50.70 & 2.80 \\
    \hline
  \end{tabular}
  \caption{Real-time performance of DeepPicar depending on the number of cores used.}
\end{table}

The performance of the DeepPicar is summarized in Table III. The DeepPicar performed better, on average, 
when it utilized more cores. With 4 cores, it was able to meet the vast majority of its deadlines, doing 
so in almost 99\% of the input frames. The platform performed the worst when using only 1 core, 
as it was unable to meet any of its deadlines.  On average, using 3 cores instead of 2 only had a 
performance increase of approximately 2 ms, so the addition of one core in that specific case offers 
relatively little improvement. One important observation is that the performance was more consistent 
when only 1 core was used. As a result, the use of multiple cores is very beneficial in terms of 
reducing the time it takes to complete inference operations, but may result in times that are more 
volatile.

\subsection{Multimodel Performance}
We also test the capability of the DeepPicar to run multiple models at the same time, and whether each 
model can successfully perform within the given deadline. Specifically, the platform is tested in the 
cases of running 2 and 4 models simultaneously. For each case, all models are allocated an equal number 
of cores. That is, 2 models are given 2 cores each and 4 models are given 1 core each.

\begin{table*}
  \centering
  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
  \hline
  \textbf{num models} & \textbf{cores} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
    misses} & \textbf{L1 miss \%} & \textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
  1 & 0,1 & 51.35 & 3.04E+10 & 4.78E+08 & 1.58 & 3.31E+09 & 3.68E+08 & 11.12\\ \hline
  2 & 0,1 & 58.03 & 3.04E+10 & 4.91E+08 & 1.61 & 3.91E+09 & 4.26E+08 & 10.88 \\ \hline
  2 & 2,3 & 56.40 & 3.04E+10 & 4.80E+08 & 1.58 & 3.88E+09 & 4.21E+08 & 10.87 \\ \hline
  \end{tabular}
  \caption{Real-time performance of the DeepPicar when 2 models are running simultaneously.}
\end{table*}

The results for the 2 model test are outlined in Table IV, and the 4 model test is summarized in Table 
V. In the the 2 model test, both of the models showed average inference time increases of around 5-7 
ms, $\sim$10\%, when compared to a baseline of 1 model running on two cores. The difference was even 
greater in the 4 model test, as each model displayed an average inference time increase of 
approximately 15 ms, $\sim$30\%, when compared to a single model running on 1 core.

The increase in inference times, however, was not the result of increased cache misses due to 
additional accesses. For all models, the number of L1 misses was always $\sim$1.6\% of all 
references. In the 2 model test, L2 cache misses remained at $\sim$11\% of all references, and each 
model in the 4 model test missed $\sim$13\% of all L2 references.

\begin{table*}
  \centering
  \begin{tabular} {| l | l | l | l | l | l | l | l | l |}
  \hline
  \textbf{num models} & \textbf{core} & \textbf{mean} & \textbf{L1 refs} & \textbf{L1 
    misses} & \textbf{L1 miss \%} & \textbf{L2 refs} & \textbf{L2 misses} & \textbf{L2 miss \%} \\ \hline
    1 & 0 & 62.48 & 2.78E+10 & 4.36E+08 & 1.57 & 2.83E+09 & 3.59E+08 & 12.68 \\ \hline
    4 & 0 & 77.90 & 2.79E+10 & 4.53E+08 & 1.63 & 3.36E+09 & 4.43E+08 & 13.19 \\ \hline
    4 & 1 & 78.89 & 2.79E+10 & 4.64E+08 & 1.67 & 3.42E+09 & 4.38E+08 & 12.82 \\ \hline
    4 & 2 & 77.81 & 2.78E+10 & 4.45E+08 & 1.60 & 3.45E+09 & 4.41E+08 & 12.77 \\ \hline
    4 & 3 & 77.87 & 2.79E+10 & 4.45E+08 & 1.60 & 3.41E+09 & 4.39E+08 & 12.88 \\ \hline
  \end{tabular}
  \caption{Real-time performance of the DeepPicar when 4 models are running simultaneously.}
\end{table*}

In all multimodel tests, it was found that a greater number of models run simultaneously resulted in 
interference that led to a noticable increase in the average inferencing time that wasn't 
attributable to a change in cache misses. Instead, the overall time increases can most likely be 
traced to an increase in cache latency. Even if the models had a consistent number of cache hits, it 
is highly probable that the contention for the cache increased the access time for each model, 
consequently increasing the time it took for the models to execute their operations.

In terms of real-time performance under a WCET of 50 ms, the DeepPicar was unsatisfactory in all 
multimodel tests. On average, the 2 model test saw models miss their deadline by 6-8 ms, and the 4 
model test was worse as each model went over their deadlines by 27-29 ms. As a result, it can be 
ascertained that, if a WCET of 50 ms is required, DeepPicar would not be able to perform as necessary.

\subsection{Summary}
We found that the DeepPicar is capable of successfully completing all necessary inferencing 
operations within a given deadline of 50 ms. Between these operations, it was discovered that the 
angle prediction time of the model was the dominating step in the processing time of a single frame. 
The other operations were found to execute in very little time, with either one taking 5 ms at most.

When running a single model, DeepPicar performed the best when it used all 4 cpu cores, as it was 
able to process a frame in 43 ms on average. This feat can still be accomplished when running a 
single model on 3 cpu cores, and is also possible when using 2 cores. The only time in which the Pi 
can't run a single model under 50 ms is when only 1 core is used. This was also the case when 
multiple models were run concurrently as no models were able to consistently complete inference 
operations in under 50 ms.

However, the DeepPicar was shown to be capable of handling other potential WCETs. If given a larger 
deadline, the capabilities of DeepPicar as an autonomous vehicle platform would be greatly increased. 
For example, if tested with a deadline of 66.67 ms, or 15 fps, the platform would have passed several 
of the experiments performed. In actuality, the DeepPicar would have only been unsatisfactory in the 
4 model test where it would have missed the deadlines by $\sim$10 ms.

@@

The most notable issue that can affect the cpu clock speed is that of the power supplied to the 
Raspberry Pi. In essence, it is necessary that the Pi be supplied with 2 Amps, as any less could 
hinder the Pi's ability to maintain a 1.2 GHz frequency. In experiments conducted with a power supply that 
only provided 1 Amp, the Pi was unable to sustain a 1.2 GHz clock speed and, instead, fluctuated 
between operating at 600 MHz and 1.2 GHz. As a result, it is necessary, or at least highly 
recommended, that the power supply used for the Raspberry Pi 3 be capable of outputting 2 Amps, 
otherwise optimal performance isn't guaranteed.

Another factor that can affect clock speed is that of the cpu's temperature. Some model operations can 
be computationally intensive, thus it is possible for the temperature of the cpu to become relatively 
high. This can be especially problematic in situations where multiple models are running 
simultaneously on the Pi. Consequently, thermal throttling may be used to decrease the clock 
speed so that the cpu temperature stays at a safe level. Thus the Raspberry Pi may not be suited 
for prolonged use, especially in cases where the workload is relatively larger, such as running multiple 
models. Rather, the Pi seems to be better suited for running in set periods, after which it is turned 
off or made idle so that the cpu is allowed time to cool down.
